```{r setup,include=FALSE}
# set the knitr options ... for everyone!
# if you unset this, then vignette build bonks. oh, joy.
#opts_knit$set(progress=TRUE)
opts_knit$set(eval.after='fig.cap')
# for a package vignette, you do want to echo.
# opts_chunk$set(echo=FALSE,warning=FALSE,message=FALSE)
opts_chunk$set(warning=FALSE,message=FALSE)
#opts_chunk$set(results="asis")
opts_chunk$set(cache=TRUE,cache.path="cache/")

#opts_chunk$set(fig.path="github_extra/figure/",dev=c("pdf","cairo_ps"))
#opts_chunk$set(fig.path="github_extra/figure/",dev=c("png","pdf"))
opts_chunk$set(fig.path="tools/figure/",dev=c("png"))
opts_chunk$set(fig.width=7,fig.height=6,dpi=100,out.width='700px',out.height='600px')

# doing this means that png files are made of figures;
# the savings is small, and it looks like shit:
#opts_chunk$set(fig.path="figure/",dev=c("png","pdf","cairo_ps"))
#opts_chunk$set(fig.width=4,fig.height=4)
# for figures? this is sweave-specific?
#opts_knit$set(eps=TRUE)

# this would be for figures:
#opts_chunk$set(out.width='.8\\textwidth')
# for text wrapping:
options(width=64,digits=2)
opts_chunk$set(size="small")
opts_chunk$set(tidy=TRUE,tidy.opts=list(width.cutoff=50,keep.blank.line=TRUE))

#epsiwal.meta <- packageDescription('epsiwal')
```

# epsiwal

[![Build Status](https://github.com/shabbychef/epsiwal/workflows/R-CMD-check/badge.svg)](https://github.com/shabbychef/epsiwal/actions)
[![codecov.io](https://codecov.io/github/shabbychef/epsiwal/graph/badge.svg?token=aUUQLyasWG)](https://app.codecov.io/github/shabbychef/epsiwal)
[![CRAN](https://www.r-pkg.org/badges/version/epsiwal)](https://cran.r-project.org/package=epsiwal)
[![Downloads](https://cranlogs.r-pkg.org/badges/epsiwal)](https://r-pkg.org/pkg/epsiwal)
[![Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/epsiwal?color=brightgreen)](https://r-pkg.org/pkg/epsiwal)

Implements conditional inference on normal variates as described in 
Lee, Sun, Sun and Taylor, "Exact Post Selection Inference, with Application to the Lasso."

-- Steven E. Pav, shabbychef@gmail.com

## Installation

This package may be installed from CRAN; the latest version may be
found on [github](https://www.github.com/shabbychef/epsiwal "epsiwal")
via devtools, or installed via [drat](https://github.com/eddelbuettel/drat "drat"):

```{r install,eval=FALSE,echo=TRUE}
# CRAN
install.packages(c('epsiwal'))
# devtools
if (require(devtools)) {
	# latest greatest
	install_github('shabbychef/epsiwal')
}
# via drat:
if (require(drat)) {
	drat:::add('shabbychef')
	install.packages('epsiwal')
}
```

# Basic Usage

First we perform some simulations under the null to show that the p-values
are uniform. We draw a normal vector with identity covariance and zero mean,
then flip the sign of each element to make them positive. We then
perform inference on the sum of the mean values. 

```{r under_null_pvals,cache=TRUE,eval=TRUE,echo=TRUE,dpi=200,out.width='600px',out.height='500px'}
library(epsiwal)
p <- 20
mu <- rep(0,p)
Sigma <- diag(p)
A <- - diag(p)
b <- rep(0,p)
eta <- rep(1,p)
Sigma_eta <- diag(Sigma)
eta_mu <- as.numeric(t(eta)%*%mu)
set.seed(1234)
pvals <- replicate(1000,{
	y <- rnorm(p,mean=mu,sd=sqrt(diag(Sigma)))
	ay <- abs(y)
	pconnorm(y=ay,A=A,b=b,eta=eta,Sigma_eta=Sigma_eta,eta_mu=eta_mu)
})
qqplot(pvals,qunif(ppoints(length(pvals))),main='p-values under procedure',
			 ylab='theoretical',xlab='empirical') 
```

```{r under_null_cis,cache=TRUE,eval=TRUE,echo=TRUE}
library(epsiwal)
p <- 20
mu <- rep(0,p)
Sigma <- diag(p)
A <- - diag(p)
b <- rep(0,p)
eta <- rep(1,p)
Sigma_eta <- diag(Sigma)
eta_mu <- as.numeric(t(eta)%*%mu)
type_I <- 0.05
set.seed(1234)
civals <- replicate(5000,{
	y <- rnorm(p,mean=mu,sd=sqrt(diag(Sigma)))
	ay <- abs(y)
	ci <- ci_connorm(y=ay,A=A,b=b,p=type_I,eta=eta,Sigma_eta=Sigma_eta)
})
cat('Empirical coverage of the',type_I,'confidence bound is around',mean(civals < eta_mu),'.\n')
```

# Maximum Likelihood Estimation

The package supports the "Selective MLE" procedure of Reid, Taylor and
Tibshirani.  This is via the `mle_connorm` function, or the associated `mle_connorm_max`.

```{r mle_comp,cache=TRUE,eval=TRUE,echo=TRUE}
library(epsiwal)
p <- 20
mu <- rep(0,p)
Sigma <- diag(p)
A <- - diag(p)
b <- rep(0,p)
eta <- rep(1,p)
Sigma_eta <- diag(Sigma)
set.seed(1234)
y <- rnorm(p,mean=mu,sd=sqrt(diag(Sigma)))
ay <- abs(y)
mval <- mle_connorm(y=ay,A=A,b=b,eta=eta,Sigma_eta=Sigma_eta)
cat('MLE is ',mval,' while true value is 0.\n')
```


# Inference on the Maximum

A common problem is to perform inference on the maximum of a normal vector.
This package provides convenience functions for this case when the vector
is assumed to have a common variance and common correlation (equicorrelation).

Suppose we observe $y_i \sim N(\mu_i, \sigma^2)$ with $\mathrm{corr}(y_i, y_j) = \rho$.
We find the maximum element $y_k$ and want to perform inference on its mean $\mu_k$,
conditional on it being the maximum.

```{r max_inference,eval=TRUE,echo=TRUE}
library(epsiwal)
set.seed(123)
n <- 8
sigma <- 1.2
rho <- 0.3
mu <- rnorm(n, mean=2)

# Generate data
Sigma <- sigma^2 * ((1 - rho) * diag(n) + rho * matrix(1, n, n))
y <- MASS::mvrnorm(1, mu, Sigma)

# Find the maximum
k <- which.max(y)
yk <- y[k]
yk1 <- y[-k]

# what is the true value?
cat('True mu_k is:', mu[k], '\n')

# Inference on mu_k
# 1. P-value for a specific mu_k
p_val <- pconnorm_max(yk, yk1, mu_k=mu[k], sigma=sigma, rho=rho)
cat('P-value for true mu_k:', p_val, '\n')

# 2. Confidence interval for mu_k
ci <- ci_connorm_max(yk, yk1, sigma=sigma, rho=rho)
cat('95% CI for mu_k:', ci, '\n')

# 3. Maximum Likelihood Estimate for mu_k
mle <- mle_connorm_max(yk, yk1, sigma=sigma, rho=rho)
cat('MLE for mu_k:', mle, '\n')
```

## See also

* The original paper, by Lee, J. D., Sun, D. L., Sun, Y. and Taylor, J. E. 
	[Exact post-selection inference, with application to the Lasso](https://arxiv.org/abs/1311.6238).
* A companion paper which considered the "Selective MLE", Reid, S., Taylor, J. and Tibshirani, R. 
	[Post-selection point and interval estimation of signal sizes in Gaussian samples](https://arxiv.org/abs/1405.3340).
* The [`PSAT` package](https://github.com/ammeir2/PSAT), which supports similar
	procedures, but is not yet on CRAN.
* The [`SelectiveInference` package](https://cran.r-project.org/package=selectiveInference),
	which implements similar inferential procedures under quadratic
	constraints, as detailed in Tibshirani, R. J., Taylor, J., Lockhart, R. and Tibshirani, R. 
	[Exact Post-Selection Inference for Sequential Regression Procedures](https://arxiv.org/abs/1401.3889).
* Two papers I wrote which needed all this functionality:
	* Pav, S. E. [Conditional inference on the asset with maximum Sharpe ratio](https://arxiv.org/abs/1906.00573).
	* Pav, S. E. [Post selection estimation of Sharpe ratios](https://arxiv.org/abs/2606.01650).

